---
title: "Week 4"
output: html_notebook
---

## Prep

```{r}
setwd("C:/programming/EDX_Data_science/2. r-programming/week 4")
```

## Q1

```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcome)
outcome[, 11] <- as.numeric(outcome[, 11])
```

```{r}
hist(outcome[, 11])
``` 
## Q2
Write a function called best that take two arguments: the 2-character abbreviated name of a state and an
outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector
with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome
in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can
be one of “heart attack”, “heart failure”, or “pneumonia”. Hospitals that do not have data on a particular
outcome should be excluded from the set of hospitals when deciding the rankings.

**Handling ties.** If there is a tie for the best hospital for a given outcome, then the hospital names should
be sorted in alphabetical order and the first hospital in that set should be chosen (i.e. if hospitals “b”, “c”,
and “f” are tied for best, then hospital “b” should be returned).

```{r}

best <- function(state, outcome) {
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  subdata <- data[,c(2, 7, 11, 17, 23)]
  colnames(subdata) <-c("name", "state", "HA", "HF", "PN")
  
  ## Check that state and outcome are valid
  if (state %in% subdata$state){
    subdata <- subdata[subdata["state"] == state,]
  }
  else {
    stop("state value not accepted")
  }
  
  if (outcome == "heart attack")
  {
    subdata <- subdata[,c(1,3)]
    subdata[, 2] <- as.numeric(subdata[, 2])
  }
  else if (outcome == "heart failure")
  {
    subdata <- subdata[,c(1,4)]
    subdata[, 2] <- as.numeric(subdata[, 2])
  }
  else if (outcome == "pneumonia")
  {
    subdata <- subdata[,c(1,5)]
    subdata[, 2] <- as.numeric(subdata[, 2])
  }
  else{stop("outcome value not accepted")}
  
  ## Return hospital name in that state with lowest 30-day death rate
  subdata <- subdata[!is.na(subdata[2]),]
  subdata[order(subdata[2], subdata[1]),][1,][[1]]
}

best("AK", "pneumonia")
  
```

## Q3

```{r}
rankhospital <- function(state, outcome, num = "best") {
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  subdata <- data[,c(2, 7, 11, 17, 23)]
  colnames(subdata) <-c("name", "state", "HA", "HF", "PN")
  
  ## Check that state and outcome are valid
  if (state %in% subdata$state){
    subdata <- subdata[subdata["state"] == state,]
  }
  else {
    stop("state value not accepted")
  }
  
  if (outcome == "heart attack")
  {
    subdata <- subdata[,c(1,3)]
    subdata[, 2] <- as.numeric(subdata[, 2])
  }
  else if (outcome == "heart failure")
  {
    subdata <- subdata[,c(1,4)]
    subdata[, 2] <- as.numeric(subdata[, 2])
  }
  else if (outcome == "pneumonia")
  {
    subdata <- subdata[,c(1,5)]
    subdata[, 2] <- as.numeric(subdata[, 2])
  }
  else{stop("outcome value not accepted")}
  
  subdata <- subdata[!is.na(subdata[2]),]
  
  if (is.character(num)){
    if (num == "best"){
      num <- 1
    } else if (num == "worst") {
      num <- nrow(subdata)
    }
    else {
      stop("num value not accepted")
    }
  } 
  
  
  ## Return hospital name in that state with lowest 30-day death rate
  subdata[order(subdata[2], subdata[1]),][num,][[1]]
}

rankhospital("NC", "heart attack", "worst")

```


## Q4

```{r}
rankall <- function(outcome, num = "best") {
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  subdata <- data[,c(2, 7, 11, 17, 23)]
  colnames(subdata) <-c("name", "state", "HA", "HF", "PN")
  
  if (outcome == "heart attack")
  {
    subdata <- subdata[,c(1,2,3)]
    subdata[, 3] <- as.numeric(subdata[, 3])
  }
  else if (outcome == "heart failure")
  {
    subdata <- subdata[,c(1,2,4)]
    subdata[, 3] <- as.numeric(subdata[, 3])
  }
  else if (outcome == "pneumonia")
  {
    subdata <- subdata[,c(1,2,5)]
    subdata[, 3] <- as.numeric(subdata[, 3])
  }
  else{stop("outcome value not accepted")}
  
  subdata <- subdata[!is.na(subdata[3]),]
  subdata <- subdata[order(subdata[2], subdata[3], subdata[1]),]
  
  if (is.character(num)){
    if (num == "best"){
      num <- 1
    } else if (num == "worst") {
      num <- "worst"
    }
    else {
      stop("num value not accepted")
    }
  } 
  
  outdf <- aggregate(subdata, by=list(subdata$state), function(x){
    if (num == "worst") {
      num <- length(x)
      x[num]
    } else {
      x[num]
    }
  })
  
  outdf <- outdf[,c("state", "name")]
  
  colnames(outdf) <- c("state", "hospital")
  outdf[,c(2,1)]
}

r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)
```